Automating Political Bias Prediction

### 2016-05-22 manuscript/rebuttal.txt

Q: d is the size of the dictionary [...] union of the words [...] in the corpus and the words [...] in the sentiment dictionary
A: Yes

Q: the sentiment vector s includes all sentiment-related words, both positive and negative.
A: It contains the sentiment polarity values associated with the key words.

Q: What is N in Table 1?
A: Number of data points, the revised manuscript includes this now.

Q: [...] difference in performance between the 17th and 18th bundestag
A: one is a 5 class the other a 4 class problem, also the 18th Bundestag is still running, so the amount of data is different.

Q: [...] performance [...] of political views is not [...] high. [...] based on what this may be considered relevant and/or useful
A: Even if the results are not perfect the model can be still highly relevant for assistive technology e.g. for speeding up human annotations in active learning scenarios.

Q: [...] experiment on correlations between words and parties raises
A: The experiments on word correlations and sentiment correlations are separate and independent experiments and discussed independently.

Q: [remark on table 9]
A: The notion of sentiment used in the manuscript has polarity.

### 2016-05-23 manuscript/rebuttal.txt

I would like to thank all three reviewers for their careful review and helpful feedback.
Especially I would like to thank reviewer 2 for providing valuable references of related work. I found these both highly relevant and interesting and I do wish I would have found them before. I read and added those and others I found meanwhile to the bibliography. Also I added a discussion of the related work to a revised manuscript and pointed out common results as well as new aspects in my results that I couldn’t find in those references.

I do hope that the reviewers would consider taking this into account and I hope that it would be possible to add these added references and their discussion to the camera ready version of the manuscript?

@@

I would like to address first the three main concerns of the reviewers and then answer some specific concerns in more detail.

1) Missing related work:
I did extensive literature research, compared and discussed as many studies I could cover and added them to the revised manuscript.

2) The analysis only covers german.
That is mostly due to the availability of data - but I will definitely look into more languages asap.

3) Low and badly explained accuracy of predictions:
I compared the results with previous studies (Yu et al, "Classifying party affiliation from political speech", 2008 and Hirst et al, "Text to Ideology or Text to Party Status?", 2014).
My results are on par with or better than those reported in previous studies (on comparable data sources). Importantly also the explanations and empirical tests thereof provided in my submitted manuscript were in line with and complemented the findings in those studies:

a) the results of binary party classification (see Yu et al, 2008) was comparable (accuracies between 0.8 and 0.87) to the accuracy in the present study (0.85 for the 17th Bundestag)
b) the accuracy in the five class classification of the 17th Bundestag (0.63) is slightly better than the only comparable study I found: Hirst et al, 2014 obtained a classification accuracy of 0.61 on a five class problem of political affiliation prediction in the European parliament.
c) As demonstrated in the submission the main sources of poor accuracy are the out-of-domain generalization challenge and the short length of the texts. Both effects were described before in Yu et al, 2008 and Hirst et al 2014, respectively, in similar (but not the same) settings.

Also it would be worth noting that even if the accuracies are (as those of comparable studies) are low, they nonetheless are on average above chance. While these results do not allow for usage in production systems for classification, it is well possible to use such a system as assistive technology for human annotators in an active learning setting.

DETAILED COMMENTS

REVIEWER 2
Q: [...] other classifiers
A: I did try others (random forests, kernel based methods); results on in-domain data were slightly better but results on out of domain data were significantly worse (for all regularizations i tried).

Q: What about stemming of words?
A: I did include this in earlier hyperparameter optimizations, but it did not improve results, so i excluded it for brevity.

Q: [sentence level too fine, topic too coarse ]
A: I tried paragraphs for manifestoproject data earlier, results were between sentence and topic performance. Hirst et al, 2014 show a comprehensive analysis of text length and accuracy, though.

@@

REVIEWER 3

Q: each sentence [...] associated to many labels?
A: No.

Q: [...] 3-fold cross-validation sufficent [...]
A: [...] A single random split into train/hyperparameter-optimization/evaluation is common practice (see e.g. Netflix prize); I did 5 fold nested CV, results were the same.

@@

Q: the sentiment vector s includes all sentiment-related words, both positive and negative.
A: It contains the sentiment polarity values associated with the key words.

Q: What is N in Table 1?
A: Number of data points, the revised manuscript includes this now.

Q: [...] difference in performance between the 17th and 18th bundestag
A: one is a 5 class the other a 4 class problem, also the 18th Bundestag is still running, so the amount of data is different.

Q: [...] experiment on correlations between words and parties raises
A: The experiments on word correlations and sentiment correlations are separate and independent experiments and are discussed independently.

### 2016-05-23 manuscript/political_bias_prediction.tex
shortened reply, polished ms text

In the following \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results and future research directions.

\section{Related Work}\label{sec:related}
Throughout the last years automated content analyses for political texts content have been conducted on a variety of text data sources (parliament data \, blogs, tweets, news articles, party manifestos) with a variety of methods, including sentiment analysis, stylistic analyses, standard bag-of-word (BOW) text feature classifiers and more advanced natural language processing tools. 
While a complete overview is beyond the scope of this work, the following paragraphs list similarities and differences between this study and previous work. For a more complete overview we refer the reader to \cite{Grimmer2013,Kaal2014}.

A similar approach to the one presented here was taken in \cite{Yu2008}. The authors extracted BOW feature vectors and applied linear classifiers to predict political party affiliation of US congress speeches. They used data from the two chambers of the US congress, House and Senat, in order to investigate generalization performance of a classifier trained on one data source and tested on another. They found that generalization of the classifier as quantified by cross-validation was in some cases strongly affected by the domain. Generalization was also affected by the time difference between the political speeches used for training and those used for testing. 
%
Other work has focused on developing dedicated methods for predicting political bias, what is often referred to as {\em scaling} in these studies; two have become particularly popular, WordFish \cite{Slapin08ascaling} and WordScores \cite{Laver2003}. Both methods have been applied in many studies and their strengths and weaknesses compared, which led to several improvements, see e.g. \cite{Lowe09scalingpolicy}. Validation is done by comparing parameter estimates of the different models amongst each other or with human expert annotations. Models are usually not evaluated by testing them on held-out data, which renders comparisons with this work or other work such as \cite{Yu2008} difficult. In general these approaches have been very valuable for a posteriori analysis of historical data but do not seem to be used as much for analyses of new data in a predictive analytics setting. 
%
Also texts from social network data were investigated in the context of automated political text analysis in the 2008 U.S. presidential election \cite{Gryc2014}. The study highlighted the importance of features such as sentiment or social network structure but the results also show that standard BOW features combined with linear classifiers yield competitive results for sentiment classification on political tweets. 
%
Finally Hirst et al conducted a large number of experiments that offer an interesting comparison to the present study both in terms of methodology but also with respect to their results \cite{Hirst2014}. The authors show that a classifier trained on parliament speeches uses language elements of defense and attack to classify speeches, rather than ideological vocabulary. They also argue that emotional content plays an important role in automatic analysis of political texts. Furthermore their results show a clear depencency between length of a political text and the accuracy with which it can be classified correctly.

Taken together, there is a large body of literature on an ever expanding field in which scientists from quantitative empirical disciplines as well as political science experts collaborate on the challenging topic of automated analysis of political texts. However except for few exceptions most previous work has focused to a large extent on binary classification\footnote{Many parliaments only have two parties and many studies chose binary classification schemes within manually defined topics and generic schemes such as conservative vs liberal or left vs right.} or to assignment of a one dimensional policy position (mostly left vs right). Often this is a too coarse approach to take into account subtle differences in political policies. This work focuses on more fine grained political view prediction: for one the case of the german parliament is more diverse than two parliament systems, allowing for a distinction between more policies, and second the political view labels considered are more fine grained than in previous studies. Another important contribution of this work is that many existing studies are primarily concerned with a posteriori analysis of historical data. This work aims at prediction of political bias on out-of-domain data with a focus on the practical application on new data, for which a prototypical web application is provided. The experiments follow the established principles of cross validation to estimate out-of-domain generalization and complement the work of \cite{Yu2008} and \cite{Hirst2014} by evaluating data from the german parliament. We confirm and extend some of their findings on why and when generalization performance is impacted negatively.

@@

\paragraph{Government membership prediction} Next to the party affiliation labels also government membership labels were used to train models that predict whether or not a text is from a party that belonged to a governing coalition of the Bundestag. In \autoref{tab:results_binary_17} and \autoref{tab:results_binary_18} the results are shown for the 17th and the 18th Bundestag, respectively. While the in-domain evaluation precision and recall values reach values close to 0.9, the out-of-domain evaluation drops again to values between 0.6 and 0.7. This is in line with the results on binary classification of political bias in the Canadian parliament \cite{Yu2008}. The authors report classification accuracy of up to 0.87, the accuracy in the 17th Bundestag was While topic-level predictions were not performed in this binary setting, the party affiliation results in \autoref{tab:results_topic} suggest that a similar increase in out-of-domain prediction accuracy could be achieved when aggregating texts to longer segments.

@@

\section{Conclusions, Limitations and Outlook}\label{sec:conclusion}
This study presents a simple approach for automated political bias prediction. The results of these experiments show that automated political bias prediction is possible with above chance accuracy in some cases.

Some aspects of these analyses could be interesting for social science researchers and three of these are highlighted in this study.
First the misclassifications of a model can be related to the changes in policy of a party. Such analyses could be helpful to quantitatively investigate a change in policy. Second analysing the word-party correlations shows that some discriminative words can be related to the political views of a party; this allows for validation of the models by human experts. Third when correlating the sentiment of a speech with proxies of political power there is a strong positive correlation between political power and positive sentiment. This supports the intuition that if we read a text that has a positive opinion on what the party in power does the author of the text is probably affiliated with those in power. While such an insight in itself might seem not very surprising this quantitatifiable link between power and sentiment could be useful nonetheless: Sentiment analysis is a rather domain independent measure, it can be easily automated and scaled up to massive amounts of text data. Combining sentiment features with other measures of political bias could potentially help to alleviate some of the domain-adaptation problems encountered when applying models trained on parliament data to data from other domains.

One of the main limitations of an automated political bias prediction system is the availability of training data. Most training data sets that are publicly available have an inherent bias as they are sampled from a different domain. This study tried to quantify the impact of this, but not in all settings out-of-domain data was available to quantify generalisation performance across text domains. 
For the cases where evaluation data from two domains was available there was a pronounced drop in performance between in-domain evaluation set and out-of-domain evaluation set. This effect was reported previously, see e.g. \cite{Yu2008}. Also the finding that shorter texts are more difficult to classify than longer texts is in line with previous studies \cite{Hirst2014}. However when considering texts of sufficient length (for instance by aggregating all texts of a given political topic, see \autoref{tab:results_topic}) the classification performance became better and in some cases reliable predictions can be obtained even beyond the training text domain.

>>>>

In the following \autoref{sec:related} briefly surveys some related work, thereafter \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results and future research directions.

\section{Related Work}\label{sec:related}
Throughout the last years automated content analyses for political texts have been conducted on a variety of text data sources (parliament data \, blogs, tweets, news articles, party manifestos) with a variety of methods, including sentiment analysis, stylistic analyses, standard bag-of-word (BOW) text feature classifiers and more advanced natural language processing tools. 
While a complete overview is beyond the scope of this work, the following paragraphs list similarities and differences between this study and previous work. For a more complete overview we refer the reader to \cite{Grimmer2013,Kaal2014}.

A similar approach to the one presented here was taken in \cite{Yu2008}. The authors extracted BOW feature vectors and applied linear classifiers to predict political party affiliation of US congress speeches. They used data from the two chambers of the US congress, House and Senat, in order to investigate generalization performance of a classifier trained on one domain and tested on another. They found that accuracies of the model when trained on one domain and test on another was significantly decreased. Generalization was also affected by the time difference between the political speeches used for training and those used for testing.

Other work has focused on developing dedicated methods for predicting political bias such as for instance two methods known as WordFish \cite{Slapin08ascaling} and WordScores \cite{Laver2003}, respectively, and improved versions thereof, see e.g. \cite{Lowe09scalingpolicy}. Validation in these studies is often done by comparing parameter estimates of the different models amongst each other or with human expert annotations. These models are often not evaluated on held-out data or with respect to the same type of labels used to train or partition the data, which renders comparisons with this work or other work such as \cite{Yu2008} difficult. These approaches have been very valuable for a posteriori analysis of historical data but do not seem to be used as much for analyses of new data in a predictive analytics setting. 
%
Also texts from social network data were investigated in the context of automated political text analysis in the 2008 U.S. presidential election \cite{Gryc2014}. The study highlighted the importance of features such as sentiment or social network structure but the results also show that standard BOW features combined with linear classifiers yield competitive results for sentiment classification on political tweets.

Finally Hirst et al conducted a large number of experiments in the Canadian parliament and the European parliament that offer an interesting comparison to the present study both in terms of methodology but also with respect to their results \cite{Hirst2014}. The authors show that a classifier trained on parliament speeches uses language elements of defense and attack to classify speeches, rather than ideological vocabulary. They also argue that emotional content plays an important role in automatic analysis of political texts. Furthermore their results show a clear depencency between length of a political text and the accuracy with which it can be classified correctly.

Taken together, there is a large body of literature in this expanding field in which scientists from quantitative empirical disciplines as well as political science experts collaborate on the challenging topic of automated analysis of political texts. However except for few exceptions most previous work has focused to a large extent on binary classification\footnote{Many parliaments only have two parties and many studies chose binary classification schemes within manually defined topics and generic schemes such as conservative vs liberal or left vs right.} or to assignment of a one dimensional policy position (mostly left vs right). Often this is a too coarse approach to take into account subtle differences in political policies. This work focuses on more fine grained political view prediction: for one the case of the german parliament is more diverse than two parliament systems, allowing for a distinction between more policies; second the political view labels considered are more fine grained than in previous studies. Another important contribution of this work is that many existing studies are primarily concerned with a posteriori analysis of historical data. This work aims at prediction of political bias on out-of-domain data with a focus on the practical application of the model on new data, for which a prototypical web application is provided. The experiments estimate out-of-domain generalization and thus complement the work of \cite{Yu2008, Hirst2014} with results from data of the german parliament. In particular the results of this study confirm and extend some of their findings on why and when generalization performance is impacted negatively.

@@

\paragraph{Government membership prediction} Next to the party affiliation labels also government membership labels were used to train models that predict whether or not a text is from a party that belonged to a governing coalition of the Bundestag. In \autoref{tab:results_binary_17} and \autoref{tab:results_binary_18} the results are shown for the 17th and the 18th Bundestag, respectively. While the in-domain evaluation precision and recall values reach values close to 0.9, the out-of-domain evaluation drops again to values between 0.6 and 0.7. This is in line with the results on binary classification of political bias in the Canadian parliament \cite{Yu2008}. The authors report classification accuracies between 0.8 and 0.87, the accuracy in the 17th Bundestag was 0.85. While topic-level predictions were not performed in this binary setting, the party affiliation results in \autoref{tab:results_topic} suggest that a similar increase in out-of-domain prediction accuracy could be achieved when aggregating texts to longer segments.

@@

\section{Conclusions, Limitations and Outlook}\label{sec:conclusion}
This study presents a simple approach for automated political bias prediction. The results of these experiments show that automated political bias prediction is possible with above chance accuracy in some cases. It is worth noting that even if the accuracies are (as those of comparable studies \cite{Yu2008, Hirst2014}) are not perfect, they nonetheless are on average above chance. While these results do not allow for usage in production systems for classification, it is well possible to use such a system as assistive technology for human annotators in an active learning setting.

One of the main limiting factors of an automated political bias prediction system is the availability of training data. Most training data sets that are publicly available have an inherent bias as they are sampled from a different domain. This study tried to quantify the impact of this, but not in all settings out of domain data was available to quantify generalisation performance across text domains. 
For the cases where evaluation data from two domains was available there was a pronounced drop in performance between in domain evaluation set and out of domain evaluation set. This effect was reported previously for similar data, see e.g. \cite{Yu2008}. Also the finding that shorter texts are more difficult to classify than longer texts is in line with previous studies \cite{Hirst2014}. However when considering texts of sufficient length (for instance by aggregating all texts of a given political topic, see \autoref{tab:results_topic}) the classification performance became better; in some cases reliable predictions can be obtained even beyond the training text domain.

Some aspects of these analyses could be interesting for social science researchers; three of these are highlighted here.
First the misclassifications of a model can be related to the changes in policy of a party. Such analyses could be helpful to quantitatively investigate a change in policy. Second analysing the word-party correlations shows that some discriminative words can be related to the political views of a party; this allows for validation of the models by human experts. Third when correlating the sentiment of a speech with proxies of political power there is a strong positive correlation between political power and positive sentiment. While such an insight in itself might seem not very surprising this quantifiable link between power and sentiment could be useful nonetheless: Sentiment analysis is a rather domain independent measure, it can be easily automated and scaled up to massive amounts of text data. Combining sentiment features with other measures of political bias could potentially help to alleviate some of the domain-adaptation problems encountered when applying models trained on parliament data to data from other domains. \\
